---
title: "Fucntional Annotation - Cell Signaling"
author: "Joshua G. Medina-Feliciano"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# CHANGE TO WHAT CORRESPONDS TO YOU
knitr::opts_knit$set(root.dir = '01-DEVNEURO-LAB/34-scRNA-seq/08-SCRNA-HGLABv4/')
```

# 1.0 Loading libraries

_Check if any of this libraries are required_

```{r}

# library(Seurat)
# library(ggplot2)
# library(dplyr)
# library(DESeq2)
# library(SoupX)
# library(hdf5r)
# library(scCustomize)
# library(DropletUtils)
# library(cowplot)
# library(metap)
# library(scater)
# library(TSCAN)
# library(readr)
# library(RColorBrewer)
# library(patchwork)
# library(tidyverse)
# library(ggnewscale)
# library(reshape2)
# library(pcaMethods)
# library(velocyto.R)
# library(SeuratWrappers)
# library(SeuratDisk)
# library(AnnotationHub)
# library(clusterProfiler)
# library(DOSE)
# library(org.Hs.eg.db)
# library(decoupleR)

```


- Note: I don't think all of them are required. But just in case.
```{r}

library(Seurat)
library(ComplexHeatmap)
library(presto)
library(CellChat)
library(NMF)

```

- Reference to package manual and tutorial:
  - http://www.cellchat.org
  - https://github.com/jinworks/CellChat
  - Article: https://www.nature.com/articles/s41467-021-21246-9

# 1.1 Load RDS of Seurat object

```{r}

#all_tissues_seurat <- readRDS(file = "00_DATA/sobj_tissue_comb_08res.v1.rds")

```

# 2. Prepare reference files

## 2.1 Import human Uniprot metadata
```{r}

human_uniprot.metadata <- read_delim(
  file = "00_DATA/UP000005640_9606.hsapUniprot.metadata.tsv", 
  delim = "\t", 
  col_names = TRUE)


head(human_uniprot.metadata)

```

## 2.2 Import features table for Humans

```{r}

human_ncbi.metadata <- read_delim(
  file = "00_DATA/GCF_000001405.40_GRCh38.p14_feature_table.txt", 
  delim = "\t", 
  col_names = TRUE)

head(human_ncbi.metadata)

```

```{r}

human_ncbi.metadata_subset <- human_ncbi.metadata %>% filter(class == "with_protein") %>% dplyr::select(product_accession, GeneID)

head(human_ncbi.metadata_subset)

```


## 2.4 Import Human BLASTp results

```{r}

hsap_blastp.df <- read_delim(
  file = "00_REFS/02_HUMAN_NCBI/augustus_hints_w_utr.v4.human_ncbi.map.tsv", 
  delim = "\t", 
  col_names = FALSE)

colnames(hsap_blastp.df) <- c('qseqid', 'product_accession')
hsap_blastp.df$qseqid <- sub("\\..*$", "", hsap_blastp.df$qseqid)
head(hsap_blastp.df)

```

### 2.4.1 Creating mapping data frame

```{r}

hsap_blastp.df <- hsap_blastp.df[!duplicated(hsap_blastp.df$qseqid), ]

hsap_mapping_df_tmp <- inner_join(hsap_blastp.df, human_ncbi.metadata_subset, by = "product_accession")
hasap_geneID_mapping_df <- hsap_mapping_df_tmp %>% dplyr::select(qseqid, GeneID)

head(hasap_geneID_mapping_df)

```

# 3.0 Setting up data for CellChat

```{r}

reg_tissues <- all_tissues_seurat

# Make sure yuou are RNA assay is active and not the integrated.
DefaultAssay(reg_tissues) <- "RNA"
dim(reg_tissues)

```

## 3.1 Get count matrix

```{r}

reg_tissues_mtx <- as.matrix(reg_tissues@assays$RNA@data)
head(reg_tissues_mtx[1:20,1:20])

```

*Creating Mapping Table*

```{r}

human_ncbi.metadata_symbols_subset <- human_ncbi.metadata %>% filter(class == "with_protein") %>% dplyr::select(product_accession, symbol)

head(human_ncbi.metadata_symbols_subset)


```

```{r}
hsap_symbol_mapping_df_tmp <- inner_join(hsap_blastp.df, human_ncbi.metadata_symbols_subset, by = "product_accession")
hasap_symbol_mapping_df <- hsap_symbol_mapping_df_tmp %>% dplyr::select(symbol, qseqid)

head(hasap_symbol_mapping_df)
dim(hasap_symbol_mapping_df)
```

*Filter matrix for genes that only have human match*
```{r}

reg_tissues_mtx <- reg_tissues_mtx[hasap_symbol_mapping_df$qseqid,]
reg_tissues_mtx[1:10,1:20]
```

*Change rownames of the matrix*
```{r}
rownames(reg_tissues_mtx) <- hasap_symbol_mapping_df$symbol
storage.mode(reg_tissues_mtx) <- "integer"
reg_tissues_mtx[1:10,1:20]
```

*Create seurat object*

```{r}

hsap_sobj <- CreateSeuratObject(counts = reg_tissues_mtx, meta.data = reg_tissues@meta.data, )
dim(hsap_sobj)

hsap_sobj[["umap"]] <- reg_tissues[["umap"]]

Idents(hsap_sobj) <- Idents(reg_tissues)
 
 # Rename identity classes
hsap_sobj <- RenameIdents(object = hsap_sobj, `0` = "C0", `1` = "C1", `2` = "C2",
                          `3` = "C3", `4` = "C4", `5` = "C5", `6` = "C6",
                          `7` = "C7", `8` = "C8", `9` = "C9",
                          `10` = "C10", `11` = "C11", `12` = "C12")

```

- Check that everything looks good.

```{r}
# Confirm that everything is in order.
#library(ggh4x)

axis <- ggh4x::guide_axis_truncated(
  trunc_lower = unit(0, "npc"),
  trunc_upper = unit(2, "cm")
)


# Check that identities names are correct...

DimPlot(hsap_sobj, reduction = "umap", label=F) + 
  guides(x = axis, y = axis) +
  theme(axis.line = element_line(arrow = arrow(length = unit(0.2, "cm"), type="closed")),
        text = element_text(size = 10),
        axis.title = element_text(hjust = 0),
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())
```


```{r}

# Try searching by human gene symbol....

FeaturePlot(hsap_sobj, features = "FZD7", min.cutoff = 1, 
            max.cutoff = 10, pt.size = 1, order = TRUE) &
  guides(x = axis, y = axis) &
  theme(axis.line = element_line(arrow = arrow(length = unit(0.2, "cm"), type="closed")),
        text = element_text(size = 10),
        axis.title = element_text(hjust = 0),
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(),
        legend.title=element_blank(), legend.position = "right",
        plot.title = element_text(size = 8, hjust = -0.02),
        legend.text=element_text(size=8),
        legend.key.size = unit(0.4, 'cm'),
        axis.text=element_text(size=8)) &
  scale_colour_gradientn(colours = rev(c("dark red", "red", "grey75")))

```


*Getting data for CellChat*

```{r}

cc_hsap_mtx <- hsap_sobj[["RNA"]]@data # normalized data matrix
cc_cell_labels <- Idents(hsap_sobj)
cc_meta <- data.frame(labels = labels, row.names = names(labels)) # create a dataframe of the cell labels

```

```{r}

cellChat <- createCellChat(object = hsap_sobj, group.by = "ident", assay = "RNA")

```

# Ligand-receptor Interaction DB

```{r}

CellChatDB <- CellChatDB.human # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)

```

```{r}

CellChatDB.use <- CellChatDB
cellChat@DB <- CellChatDB.use

```

```{r}

cellChat <- subsetData(cellChat)

```

```{r}

cellChat <- identifyOverExpressedGenes(cellChat)
cellChat <- identifyOverExpressedInteractions(cellChat)

```

```{r}

cellChat <- computeCommunProb(cellChat, type = "truncatedMean", trim = 0.1)

```


```{r}

cellChat <- filterCommunication(cellChat, min.cells = 10, rare.keep = T)

```

```{r}

cc_df.net <- subsetCommunication(cellChat)
#df.net <- subsetCommunication(cellchat, sources.use = c(1,2), targets.use = c(4,5))
#df.net <- subsetCommunication(cellchat, signaling = c("WNT", "TGFb"))
```


```{r}

cellChat <- computeCommunProbPathway(cellChat)

```


```{r}
cellChat <- aggregateNet(cellChat)

```


```{r}

groupSize <- as.numeric(table(cellChat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellChat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellChat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")

```


```{r}

mat <- cellChat@net$weight
par(mfrow = c(3,5), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}

```

*Visualization of specific pathways*

```{r}

pathways.show <- c("L1CAM")
# Hierarchy plot
# Here we define `vertex.receive` so that the left portion of the hierarchy plot shows signaling to fibroblast and the right portion shows signaling to immune cells 
vertex.receiver = seq(1,4) # a numeric vector. 
netVisual_aggregate(cellChat, signaling = pathways.show, vertex.receiver = vertex.receiver)
# Circle plot
par(mfrow=c(1,1))
netVisual_aggregate(cellChat, signaling = pathways.show, layout = "circle")

```


```{r}
netAnalysis_contribution(cellChat, signaling = pathways.show)
```


```{r}

# Access all the signaling pathways showing significant communications
pathways.show.all <- cellChat@netP$pathways
# check the order of cell identity to set suitable vertex.receiver
levels(cellChat@idents)
vertex.receiver = seq(1,4)
for (i in 1:length(pathways.show.all)) {
  # Visualize communication network associated with both signaling pathway and individual L-R pairs
  netVisual(cellChat, signaling = pathways.show.all[i], vertex.receiver = vertex.receiver, layout = "hierarchy")
  # Compute and visualize the contribution of each ligand-receptor pair to the overall signaling pathway
  gg <- netAnalysis_contribution(cellChat, signaling = pathways.show.all[i])
  ggsave(filename=paste0(pathways.show.all[i], "_L-R_contribution.pdf"), plot=gg, width = 3, height = 2, units = 'in', dpi = 300)
}

```


```{r}

plotGeneExpression(cellChat, signaling = "Netrin", enriched.only = TRUE, type = "dot", color.use = brewer.pal(10,"Accent"))
plotGeneExpression(cellChat, signaling = "Netrin", enriched.only = TRUE, type = "violin")

```


*Network Centrality Scores*

```{r}

cellChat <- netAnalysis_computeCentrality(cellChat, slot.name = "netP")

```

```{r}
netAnalysis_signalingRole_network(cellChat, signaling = pathways.show, width = 8, height = 2.5, font.size = 10)
```

```{r}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
gg1 <- netAnalysis_signalingRole_scatter(cellChat)
gg1
```


```{r}

# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellChat, pattern = "outgoing")
ht2 <- netAnalysis_signalingRole_heatmap(cellChat, pattern = "incoming")
ht1 + ht2

```



